Good Empirics Dominate Bad Theory: Comments on Blair and Sambanis, 2020, \"Forecasting Civil Wars: Theory and Structure in an Age of 'Big Data' and Machine Learning\", *JCR*

### 2020-06-04 paper.Rmd
end of day two
did a bit more...

>>>>

B&S claim (page 3) to show that a model informed by procedural theories of escalation and de-escalation can predict the onset of civil wars ``remarkable accurately.'' Indeed they argue that this so-called theoretical model outperforms four other ``more mechanical'' alternatives. Second, they clam that the integration of structure with process are better over short forecasting windows. Third they preregisted the list of thirty countries which have the highest risk of civil war onset. They claim that such prospective predictions are rare in the literature when in fact they have been routine for many years with several prominent projects. B&S claim to be unique in assessing these forecasts. A qualitative analysis of their predictions allows them to conclude that their model is strong and a more precise test is undertaken. We will return to their analysis later, after correcting the procedural mistakes we found in their research process.

Before proceeding, we quote B&S (page 24):
\begin{quote}
Our theoretically driven model generates accurate forecasts, with base specification
AUCs of .82 and .85 over one- and six-month windows, respectively, and AUCs
as high as .92 in other specifications. Our model also consistently and sometimes
dramatically outperforms the alternatives we test...Cederman and Weidmann (2017, 476) argue that ``the hope that
big data will somehow yield valid forecasts through theory-free ‘brute force’ is
misplaced in the area of political violence.'' Our results lend some credence to this
claim.
\end{quote}
